\mytitle

### 2015-08-02 MSc_diss.tex
Rewriting literature review mess

@@

@@

\subsubsection*{Biomass burning source}
\citep{Newland2013}
[Alkyl nitrates also have a biomass burning source (Friedli et al., 2001; Simpson et al., 2002). Simpson et al. (2002) report emissions of C1 – C4 alkyl nitrates from savanna fires in Australia, with methyl nitrate being primarily emitted during the ’flaming’ stage and ethyl – butyl nitrates mainly emitted during the ‘smouldering’ stage. They estimate global emissions from this source to be on the order of 8 Gg yr-1]

@@

\subsection{Mean AN concentrations}
mean ANs concentrations: \citep{Reeves2007}, \citep{Roberts1998}
\citep{Atherton1989}
Alkyl nitrate concentrations of 0.8-37 pptv (2-butyl and 2- and 3-pentyl nitrates) in remote marine environments [Atlas, 1988] and approximately 2 ppbv in urban environments [Hanst et al., 1982] have been measured.
\citep{Chuck2002}
The atmospheric mole fraction of MeONO2 ranged from 1.2 to 43 pptv (Fig. 2,Aand B) and 0.8 to 10 pptv for EtONO2 (Fig. 2, D and E). These values lie within the range of concentrations reported in the literature for MeONO2 (0.2 to 50 pptv) and EtONO2 (0.3 to 12.2 pptv) in the remote marine troposphere (2, 8, 9).
\citep{Blake2003}
Pacific Ocean methyl nitrate 50 pptv

@@

\subsubsection{Modelling studies}
A huge number of modelling studies has been done about alkyl nitrates.

>>>>

\citep{Neuman2012}
Ozone production efficiency (OPE), which can be determined observationally from the ratio of O3 to NOx oxidation products, is used regularly as a metric to assess O3 photochemistry [e.g., Trainer et al., 1993; Neuman et al.,2009]

@@

\citep{Simpson2006} Pearl River Delta
Mixing ratios of seven C1–C5 alkyl nitrates (RONO2) were measured during a 16-month study (August 2001–December
2002) at Tai O, a coastal site 30km west of central Hong Kong in the Pearl River Delta, the fastest-growing industrial region in the world. The C3–C4 (rather than C1–C2) RONO2 were most abundant throughout the study, showing the importance of photochemical (rather than marine) RONO2 production in the sampled air. A lack of methyl nitrate (MeONO2) enhancement during summer, when the prevailing winds are from the ocean, indicates that the South China Sea is not a region of strong RONO2 emissions. By contrast, MeONO2 levels during pollution episodes (up to 25 parts per trillion by volume (pptv)) were the highest that our group has recorded during urban photochemical RONO2 production, as opposed to marine emissions or biomass burning. The highest summed RONO2 level of the study (204 pptv) was measured in the afternoon of 7 November 2002, during an intense pollution episode that captured the highest ozone (O3) level ever recorded in Hong Kong (203 ppbv). During pollution episodes, the average ratio of O3 to summed RONO2 was
roughly 1000:1 in freshly polluted air (ethyne/CO?3–5 pptv/ppbv) and 500:1 in very freshly polluted air (ethyne/ CO?6–8 pptv/ppbv). Ozone and RONO2 share a common photochemical source, and their good correlation in pollution plumes shows that RONO2 can be used as a tracer of photochemical O3 production. Even MeONO2 showed similar
diurnal variations as the C2–C5 RONO2, indicating a strong photochemical source despite its very slow photochemical production from methane oxidation. The decomposition of longer-chain alkoxy radicals also does not explain the high MeONO2 levels, and rough calculations show that methoxy radical reaction with NO2 appears to be a viable alternate pathway for MeONO2 production in polluted atmospheres, though further measurements and modeling are required to confirm this mechanism.
\citep{Comment}
It has generally been assumed that reaction (3) dominates over reaction (2) and that production of CH3ONO2 would be small. Simpson et al. (2006) suggest that model studies would be useful to test their theory. In this short comment on their paper, some model simulations are presented that confirm their hypothesis.The model simulations have been carried out using a box model running the CRI v1.0 mechanism (Jenkin et al., 2002), which is a reduced scheme of the Master Chemical Mechanism (Jenkin et al., 1997). The box model was integrated using emissions of NOx and VOC taken from the UK National emissions inventory and would be repre- sentative therefore of European rather than Asian emission profiles but are still instructive. The model is designed to simulate regional photochemical episodes and runs during European summer time (mid June) under moderately high temperature (approximately 25 1C) and relative humidity (70\%). The concentration of CH3ONO2 from the two routes, reaction (1b) and reaction (2), have been compared as a function of NOx and are summarised in Fig. 1. It is clear that above about 10 ppb of NO2 channel (2) becomes important as a source of CH3ONO2 and at around 35 ppb starts to become the dominant channel in the simulations performed.

\citep{Stroud2001} plane+colorado, plane+North Atlantic
During both campaigns alkyl nitrates were measured aboard the NOAA Orion WP-3 aircraft with a vertical range from the boundary layer to the midtroposphere (7 km) [Hiibler et al.,1998]. During STERAO 1996 the aircraft was based at Buckley Air Force Base in Denver, Colorado. Flights were conducted from June 13, 1996, to July 16, 1996, over the northeastern plains of Colorado to study how thunderstorms affect the chemical composition of the atmosphere.
Alkyl nitrates were measured over the northeastern plains of Colorado during the Stratospheric Tropospheric Exchange: Radiation, Aerosols and Ozone (STERAO) 1996 campaign and over the North Atlantic during the North Atlantic Regional Experiment (NARE) 1997. A comparison of alkyl nitrate to hydrocarbon ratios with a sequential reaction scheme model suggests our understanding of C4-Cs alkyl nitrate photochemistry is adequate to reproduce the observations. Air masses with larger ratios of alkyl nitrate to hydrocarbons were observed during NARE 1997 than STERAO 1996, consistent with the NARE 1997 measurements being in the more remote North Atlantic. These results suggest that information related to the photochemical age of an air  mass can be derived from alkyl nitrate to hydrocarbon ratios.

\citep{Worton2010} TORCH
Alkyl nitrates (C1eC5) were measured at two sites (near urban and rural) in southeast England during the Tropospheric Organic Chemistry Experiment (TORCH). Methyl nitrate was the dominant species during both campaigns accounting for on average about one third of the total measured alkyl nitrates. High mixing ratios (>50 pptv) and variability of methyl nitrate were observed at the near urban site (TORCH1) that were not seen at the rural site (TORCH2) and which could not be explained by local photochemical production or direct emissions. The diurnal variation of methyl nitrate during TORCH1 showed a morning maximum that would be consistent with nighttime chemistry followed by transport to the surface by boundary layer dynamics. Similarly, elevated morning mixing ratios were also observed during TORCH2 although the magnitudes were much smaller. As a result, methyl nitrate could represent a tracer for nighttime chemistry seen at the ground the following day. At both campaigns, the dominant source of short chain alkyl nitrates and carbonyl precursor radicals (?C4) were from decomposition of larger compounds. The magnitude of the source increased with decreasing carbon number consistent with increasing total precursor abundance. Non-photochemical emissions of acetaldehyde and acetone could not be accounted for by automobile exhaust emissions alone and indicated that other direct sources are likely important in this environment.
Alkyl nitrate measurements were made at two sites (near urban
and rural) as part of the TORCH experiment. Observed mixing ratios
were higher (factor of ?2) during TORCH1, especially during the identified heatwave period. Methyl nitrate accounted for w30\% of the total measured C1eC5 species and during TORCH1 was often present at mixing ratios > 50 pptv. The diurnal profile of methyl nitrate showed an early morning maximum that we attributed to nighttime production from NO3 chemistry in the residual layer followed by transportation to the surface by boundary layer dynamics. Further work is needed to evaluate our hypothesis of a nighttime source of methyl nitrate but if proven, methyl nitrate could represent an interesting tracer of nighttime chemistry aloft. The predominant photochemical source of the short chain alkyl nitrates and carbonyl compounds was from the photochemical oxidation and decomposition of longer chain compounds as opposed to the oxidation of the parent hydrocarbons similar to previous studies. The magnitude of this source shows a preference for the shortest chain species consistent with the increasing number of precursor compounds identified in the MCM. Primary sources of acetaldehyde and acetone were also significant during TORCH1 and could not be accounted for by automobile exhaust emissions alone. Further work to identify and quantify emissions from additional direct sources is necessary if appropriate control strategies are to be implemented.

\citep{deKock1994}
Air samples were collected at Cape Recife (34 ° 1' S 25 ° 40'E) situated on the south-east coast of Southern Afi'ica (Figure 1). The suburban, urban and industrial areas of Port Elizabeth lie approximately 10-40 kin north to north-west of the sampling site.
Mean sumANs 12.7.
Variations in absolute and relative concentrations of the alkyl nitrates could be related to changes in meteorological conditions. It is considered that the differences observed in chemical composition may represent different sources of trace gas emissions or different degrees of photochemical processing of similar sources. The meas~ed levels of the alkyl nitrates suggest long-range transport to play a major role in regulating the distribution of alkyl nitrates in the troposphere.

\citep{Beyersdorf2010}
South Pole, other Antarctic, ANs only

@@

\citep{Reeves2007} ocean+urban+fire
This paper is based on alkyl nitrate measurements made over the North Atlantic as part of the International Consortium for Research on Atmospheric Transport and Transformation (ICARTT). The focus is on the analysis of air samples collected on the UK BAe-146 aircraft during the Intercontinental Transport of Ozone and Precursors (ITOP) project, but air samples collected on board the NASA DC-8 and NOAA WP-3D aircraft as part of a Lagrangian experiment are also used. The ratios between the alkyl nitrates and their parent hydrocarbons are compared with those expected from chemical theory. Further, a box model is run to investigate the temporal evolution of the alkyl nitrates in three Lagrangian case studies and compared to observations.
The general agreement between the alkyl nitrate data and photochemical theory suggests that during the first few days of transport from the source region, photochemical production of alkyl nitrates, and thus ozone, had taken place.
Concentrations of methyl, ethyl, 1-propyl, 2-propyl, 2-butyl, 2-methyl-3-butyl and the sum of 2- and 3-pentyl nitrates (hereafter referred to as 2 + 3-pentyl nitrate) are presented.
The alkyl nitrate concentrations observed during
ITOP are typical of those found a few days downwind of precursor source regions. Although there is some evidence for an oceanic source of methyl nitrate, for the most part the air samples collected during ITOP do not appear to be strongly influenced by oceanic sources of the alkyl nitrates. Elevated concentrations were observed in air masses which have been influenced by emissions from the N.E. United States and from Alaskan fires. The large variations in ratios between the C5 and C2-C4 nitrates suggest that emission ratios for the parent hydrocarbons vary between different regions of the N.E. United States. [49] In general the alkyl nitrate data agrees well with photochemical theory. This suggests that at least during the first few days of transport from the source region photo- chemical production of alkyl nitrates had taken place. This implies that ozone production had also occurred. There also appears to be a widespread common source of ethyl nitrate and 1-propyl nitrate other than from their parent hydro- carbons. This might be from longer-chain hydrocarbons, possibly pentanes. [50] The observations in the more photochemically pro- cessed air masses are consistent with the concentration of NO dropping to a point where the rates of the alkyl nitrate production reactions no longer dominate the peroxy radical self/cross reactions. However, in these low levels of NO, where the balance between ozone production and destruc- tion is highly sensitive to NOX, the alkyl nitrates might contribute to maintaining the ozone levels in air as it undergoes long-range transport. [51] The observed patterns of the alkyl nitrates also suggest that the rates of photochemical processing in the Alaskan smoke plumes were small and that the rates of ozone formation in these air masses were low.

\citep{Simpson2003} airborne too
TRACE-P samples that did not show influence from marine alkyl nitrate sources were used to investigate photochemical alkyl nitrate evolution. Relative to 2-butyl nitrate/n-butane, the measured ratios of ethyl nitrate/ethane and 2-propyl nitrate/propane showed notable deviations from modeled values based on laboratory kinetic data, suggesting additional Asian sources of their alkyl peroxy radical precursors. By contrast, the measured ratios of 1-propyl-, 2-pentyl-, and 3-pentyl nitrate to their respective parent hydrocarbons were fairly close to modeled values.

\citep{Roberts1998} marine MBL+urban ANs
The measurements described herein were made as part of the
Organic nitrates are a class of product species that contain
North Atlantic Regional Experiment (NARE) during August and September 1993. These measurements consisted of NO, NO 2, particulate nitrate, HNO3, PAN, PPN, RONO 2, NOy, 0 3, CO, SO 2, and C2-C6 hydrocarbons at Chebogue Point. This site was ideally located to study the evolution of photochemically processed air masses in the marine boundary layer (MBL). Chebogue Point is approximately 500 km downwind from the eastern sea coast of the United States, and is often impacted by transport of polluted air from the continental source regions.
Northwest Territories, Canada, have been made in the winter to
The correlations between organic nitrates and CO or O3 have been presented by Roberts et al. [1996]. It has been established that CO is a good tracer of continentally derived pollution at this site [Parrish et al., 1993]. This continentally derived pollution contains hydrocarbons (HC) and NO x, the precursors of organic nitrates and ozone, and serves to link the sources of CO with the sources of organic nitrates and ozone. There are distinct differences in the lifetimes of CO, O3 , and organic nitrates, and the factors that they are dependent on. For these reasons, organic nitrates correlate to varying degrees with CO and O3.
Measurements of the organic nitrates; PAN, PPN, methyl-,
ethyl-, 1-propyl-, 2-propyl-, 2-butyl, and 3-pentyl nitrates were made at Chebogue Point, Nova Scotia, during the 1993 NARE campaign. The relationships of these compounds with other measures of continental pollution, for example, CO and 03, are consistent with our view of these compounds as products of HC-NO x photochemistry.
Comparisons of the ratios with that of [2-butyl nitrate]/[butane] were used to test whether the actual data fit the line predicted from rate constants, branching ratios, photolysis rates, and assumed [OH].
Abstract. Measurements of six alkyl nitrates, methyl-, ethyl-, 1-propyl-, 2-propyl-, 2-butyl, and 3-pentyl nitrates, and two peroxycarboxylic nitric anhydrides, peroxyacetic nitric anhydride
(PAN) and peroxypropionic nitric anhydride (PPN) were made
at Chebogue Point, Nova Scotia, during the 1993 North Atlantic
Regional Experiment campaign. The correlation of alkyl nitrates
(RONO2) with carbon monoxide makes it clear that the chemistry at this site is being driven by continental emissions, since RONO2 compounds are known to be by-products of HC-NOx photochemistry.
Comparison of the ratios of alkyl nitrates to their parent hydrocarbons to that of 2-butyl nitrate/butane showed significant
deviations from trends predicted from rate constants, branching
ratios, and loss rates. For ethyl nitrate this deviation is likely due to a combination of many additional pathways  o the intermediate
peroxy radical from the decomposition of larger alkoxy radicals.
The propyl nitrates showed trends that differed from one another; 2-propyl nitrate was a factor of 2 to 3 higher than predicted, while 1-propyl nitrate was a factor of 8 to 10 higher than predicted. The trend of 3-pentyl nitrate/pentane was relatively close (about
50\% lower) to the predicted trend.

\subsubsection*{Biomass burning source}
\citep{Newland2013}
[Alkyl nitrates also have a biomass burning source (Friedli et al., 2001; Simpson et al., 2002). Simpson et al. (2002) report emissions of C1 – C4 alkyl nitrates from savanna fires in Australia, with methyl nitrate being primarily emitted during the ’flaming’ stage and ethyl – butyl nitrates mainly emitted during the ‘smouldering’ stage. They estimate global emissions from this source to be on the order of 8 Gg yr-1]
\citep{Simpson2002}
We report the first observations of the emission of five C1–C4 alkyl nitrates (methyl-, ethyl-, n-propyl-, i-propyl-, and 2-butyl nitrate) from savanna burning. Average alkyl nitrate mixing ratios in the immediate vicinity of three bushfires in Northern Australia were 47–122 times higher than local background mixing ratios. These are the highest alkyl nitrate mixing ratios we have ever detected, with maximum mixing ratios exceeding 3 ppbv for methyl nitrate. Methyl nitrate dominated the alkyl nitrate emissions during the flaming stage of savanna burning, whereas C2–C4 alkyl nitrates were mostly emitted during the smoldering stage. To explain the formation of alkyl nitrates from biomass burning, we propose a reaction mechanism involving the combination of reactive radicals at high temperature. Bearing in mind the uncertainties associated with extrapolating small data sets to much larger scales, alkyl nitrate emissions from global savanna burning are estimated to be on the order of 8 Gg/yr.

@@

\subsection{Mean AN concentrations}
mean ANs concentrations: \citep{Reeves2007}, 10 ppt \citep{Roberts1998} 14 ppt
\citep{Atherton1989}
Alkyl nitrate concentrations of 0.8-37 pptv (2-butyl and 2- and 3-pentyl nitrates) in remote marine environments [Atlas, 1988] and approximately 2 ppbv in urban environments [Hanst et al., 1982] have been measured.
\citep{Chuck2002}
The atmospheric mole fraction of MeONO2 ranged from 1.2 to 43 pptv (Fig. 2,Aand B) and 0.8 to 10 pptv for EtONO2 (Fig. 2, D and E). These values lie within the range of concentrations reported in the literature for MeONO2 (0.2 to 50 pptv) and EtONO2 (0.3 to 12.2 pptv) in the remote marine troposphere (2, 8, 9).
\citep{Blake2003}
Pacific Ocean methyl nitrate 50 pptv
\citep{Neuman2012}
approx 1 ppb as difference between NOy and (NOx+PAN+HNO3) is sample closest to the plume center

@@

\subsubsection{Modelling studies}
A huge number of modelling studies has been done about alkyl nitrates. In particular, O3 and the RONO2 oxida- tion products formed from reactions between these NOx and VOC emissions are examined in detail...

\citep{Neuman2012}
Source characterization and photochemical processes in transported air masses have been analyzed by examining RONO2 formation from anthropogenic emissions of alkane precursors [Simpson et al., 2003; Madronich, 2006; Simpson et al., 2006; Reeves et al., 2007; Worton et al., 2010].
Reactive nitrogen species and volatile organic compounds (VOCs) were measured from an instrumented aircraft during daytime flights in the marine boundary layer downwind from the area of surfacing oil. A unique VOC mixture, where alkanes dominated the hydroxyl radical (OH) loss rate, was emitted into a clean marine environment, enabling a focused examination of O3 and RONO2 formation processes. In the atmospheric plume from DWH, the OH loss rate, an indicator of potential O3 formation, was large and dominated by alkanes with between 5 and 10 carbons per molecule (C5–C10). Observations showed that NOx was oxidized very rapidly with a 0.8 h lifetime, producing primarily C6–C10 RONO2 that accounted for 78\% of the reactive nitrogen enhancements in the atmospheric plume 2.5 h downwind from DWH. Both observations and calculations of RONO2 and O3 production rates show that alkane oxidation dominated O3 formation chemistry in the plume. Rapid and nearly complete oxidation of NOx to RONO2 effectively terminated O3 production, with O3 formation
yields of 6.0 plus minus 0.5 ppbv O3 per ppbv of NOx oxidized.
Eight RONO2 compounds with between one and five carbon atoms
per molecule were measured, and in a few samples the sum of C6–C8 RONO2.
NOx loss by OH + NO2 was negligible here, as demonstrated by the absence of HNO3 production in the plumes (Figure 4). Instead, the rapid NOx loss was caused by the reaction of peroxy radicals with NO.
Figure 4 illustrates temporal changes in reactive nitrogen partitioning in the DWH plume. In fresh plumes (e.g., Figure 4a), NOx accounted for over 80\% of plume NOy. In aged plumes (e.g., Figure 4b), NOx oxidation was nearly complete, but NOx, PAN, HNO3, and the sum of eight C1–C5 RONO2 accounted for only 22\% of plume NOy. This result differs dramatically from previous studies, where the sum of NOx + PAN + HNO3 accounted for over 96\% of NOy in urban and petrochemical plumes [Neuman et al., 2002; Ryerson et al., 2003] and for approximately 90\% of NOy in rural areas [Parrish et al., 1993]. In urban air masses, RONO2 often accounted for a few percent of NOy [O’Brien et al., 1997; Flocke et al.,1998; Schneider et al., 1998; Talbot et al., 2003; Ryerson et al., 2003], and up to 10–20\% of NOy in some studies [Day et al., 2003; Perring et al., 2010].
The largest contribution to the total calculated RONO2 production rate is from C6–C10 alkanes, which were responsible for 80\% of the RONO2 production.
OPE. In the DWH plume, OPE was 6.0 plus minus 0.5 ppbv/ppbv after 1 h of transport (Figure 3b), and similar to values measured in U.S. urban areas [Trainer et al., 1995; Neuman et al., 2009]. In contrast, petrochemical plumes with OH loss rates dominated by alkenes had OPEs ranging from 10 to 18 [Ryerson et al., 2003] (Figure 3b). On 10 June, O3 enhancements in the DWH plumes were less than 15 ppbv (Figure 5b), even after NOx was completely oxidized. Although the ratio of VOC to NOx was very large, O3 production was limited by OH oxidation of C5–C10 alkanes that preferentially formed RONO2 and by NOx levels less than 10 ppbv.
